Naturalness in U(1) extensions of the MSSM

### 2014-12-19 paper/e6ssm-tuning-paper.tex
added more to conclusion.  maybe it's not so shit after all, will reread later.

V_{E_6}^F&=\lambda^2|S|^2(|H_1|^2+|H_2|^2)+\lambda^2|(H_1\cdot H_2)|^2,\label{eq:E6VFterms}\\
V_{E_6}^D&=\frac{\bar{g}^2}{8}\left ( |H_2|^2-|H_1|^2\right )^2+\frac{g_2^2}{2}|H_1^\dagger H_2|^2+\frac{g_1'^2}{2}(\tilde{Q}_1|H_1|^2+\tilde{Q}_2|H_2|^2+\tilde{Q}_S|S|^2)^2,\label{eq:E6VDterms}\\
V_{E_6}^{\textrm{soft}}&=m_S^2|S|^2+m_{H_d}^2|H_1|^2+m_{H_u}^2|H_2|^2+\Big [\lambda A_\lambda S(H_1\cdot H_2)+\textrm{h.c.}\Big ].\label{eq:E6Vsoft}

@@

The quantities $D_{H_1}$, $D_{H_2}$ and $D_S$ appearing above are $U(1)'$ $D$-term contributions that are absent in the MSSM and NMSSM and are given by

@@

D_\phi\equiv \frac{g_1'^2}{2}\left ( \tilde{Q}_1v_1^2+\tilde{Q}_2v_2^2+\tilde{Q}_Ss^2\right )\tilde{Q}_\phi.

@@

As was noted in \cite{Athron:2013ipa}, the first two of these conditions may be rewritten in the form

@@

with $M_Z^2=\bar{g}^2v^2/4$, $v^2=v_1^2+v_2^2$ and $\tan\beta = v_2/v_1$ and where we have for convenience absorbed the effects of the loop corrections into the soft masses, 

@@

\tilde{m}_{H_d}^2&=m_{H_d}^2+\frac{1}{v_1}\frac{\partial \Delta V_{E_6}}{\partial v_1},\\
\tilde{m}_{H_u}^2&=m_{H_u}^2+\frac{1}{v_2}\frac{\partial \Delta V_{E_6}}{\partial v_2}.

@@

Written like so it is evident that the fine tuning contribution coming from the new $D$-terms depends both on the $U(1)'$ charges and the $Z^\prime$ mass, and that the tuning can be expected to increase with $M_{Z^\prime}$. As shall be shown below, the exact size of this tuning then depends strongly on the choice of $U(1)'$ charges, via the coefficient $d$. 

@@

\Delta_p=\left | \frac{\partial \log M_Z^2}{\partial \log p}\right |=\left | \frac{p}{M_Z^2}\frac{\partial M_Z^2}{\partial p}\right |,

@@

where $M_Z^2=\bar{g}^2v^2/4$. The coefficient $\Delta_p$ measures the fractional variation in $M_Z^2$ resulting from a given variation in the parameter $p$. The overall fine tuning is then taken to be $\Delta = \underset{i}{\max}\,\{\Delta_{p_i}\}$.\\ \\

The sensitivity coefficients $\Delta_p$ may be calculated directly from the expression for $M_Z^2$ in terms of the $p_i$ for a particular model, which leads to a so-called master formula for calculating the fine tuning. A master formula for the E$_6$SSM, obtained from the tree level scalar potential, was presented in \cite{Athron:2013ipa}. In order to derive the expression presented there, the fact that $s \gg v$ was made use of to neglect certain $O(v^2)$ terms in the EWSB conditions, greatly simplifying the final result. For the purposes of exploring a wider class of $E_6$ inspired models, we have rederived the master formula without neglecting any $O(v^2)$ terms. The more complete tree-level master formula is somewhat complicated. This is because, unlike in the MSSM, even at tree level it is not possible to solve explicitly for the VEVs $v_1$, $v_2$ in terms of the Lagrangian parameters. It may be written in the form
\Delta_p=|C|^{-1}\times \frac{|p|}{M_Z}\bigg|\sum_{q}\tilde{\Delta}_q\frac{\partial q}{\partial p}\bigg|,

@@

where the sum is over all low energy running parameters appearing in the tree level EWSB conditions, i.e. $q\in\{\lambda, A_\lambda, m_{H_d}^2,m_{H_u}^2,m_S^2,g_1,g_2,g_1'\}$. Expressions for the quantity $C$ and the $\tilde{\Delta}_q$ appearing above are given in Appendix \ref{app:masterformula}. It should be noted that the effects of $U(1)$ mixing are neglected in deriving Eq. (\ref{eq:E6masterformula}).\\ \\
However, it is well known in the MSSM that radiative corrections can significantly change (indeed, reduce) the fine tuning \cite{Cassel:2010px}. It is therefore important when studying the fine tuning to include loop corrections to the effective potential in the fine tuning measure. To do so it is most convenient to work in terms of the EWSB conditions Eq. (\ref{eq:E6EWSBConditions}), rather than Eq. (\ref{eq:E6MZequation}). The general procedure is as follows \cite{Ellwanger:2011mu}. For a model in which $m$ fields develop real VEVs (e.g. $m=2$ in the MSSM, $m=3$ in the NMSSM and in the $E_6$ models considered), we require that the $m$ minimization conditions

@@

continue to hold under an arbitrary variation in a model parameter $p\rightarrow p+\delta p$, so that the variations $\delta f_i$ satisfy

@@

Each $f_i$ is a function of the VEVs $v_j$ and $l$ running parameters $q_k$ evaluated at the scale of EWSB, $f_i=f_i(v_j,q_k)$. Thus for each $f_i$ we have that

@@

\delta f_i = \sum_{j=1}^m \frac{\partial f_i}{\partial v_j}\frac{\partial v_j}{\partial p}+\sum_{k=1}^l \frac{\partial f_i}{\partial q_k}\frac{\partial q_k}{\partial p}=0.

@@

The quantities $\frac{\partial f_i}{\partial v_j}$ are the elements of the CP-even Higgs squared mass matrix $M_h^2$ of the model before rotating into the mass eigenstate basis. When evaluated for all $n$ model parameters, the above system of equations can be concisely expressed as

@@

The quantities forming the first matrix on the right-hand side, along with $M_h^2$, are easily calculated analytically. The remaining derivatives $\partial q_k /\partial p$ must be determined using the RGEs. Once these have been obtained, it is straightforward to solve for the $\partial v_i /\partial p$. The sensitivity coefficients $\Delta_p$ are then simply linear combinations of the $\partial v_i/\partial p$ and $\partial q_k/\partial p$. Radiative corrections may be easily incorporated by including the Coleman-Weinberg potential contributions in the EWSB conditions.\\ \\
Evaluating the derivatives $\partial q_k/\partial p$ must (in general) be done by numerically integrating the two-loop RGEs. This is time consuming and presents an obstacle to doing large scans of the parameter space. For studying models defined at low energies, as we do here, we can take advantage of the fact that the amount of running involved is small. This makes it possible to use approximate analytic solutions to the RGEs that exhibit good accuracy over the range of scales considered. Given the two-loop RG equation for a parameter $q$,

@@

a Taylor series expansion of the solution may be used to obtain the parameter at the scale $Q$,

@@

Expanded to this order we obtain the leading log (LL) and next-to-LL (NLL) contributions at two-loop order. The $\mathcal{O}(t^2)$ terms not displayed above are formally of 3-loop order and are neglected. The derivative of the one-loop $\beta$ function is given by

@@

where the sum is over all running parameters appearing in $\beta_q^{(1)}$. The $\beta$ functions appearing on the right hand side of Eqs. (\ref{eq:rgeapproxsoln}) and (\ref{eq:1loopbetaderivative}) are evaluated at the input scale $M_X$, giving a simple analytic expression for the parameters at the scale of EWSB in terms of the model parameters at $M_X$. Explicit results for the relevant series expansions in the MSSM and $E_6$ models are presented in Appendix \ref{app:rges}.

@@

Using the approach outlined above, we are able to scan the low energy parameter space of the MSSM and E$_6$SSM and calculate the fine tuning in each. To do so, we implemented the above expressions for computing the fine tuning in a modified version of the E$_6$SSM spectrum generator that was used in \cite{Athron:2013ipa}. This code implemented two-loop RGEs for all parameters except the soft scalar masses. In order to properly include the fine tuning impact of the $SU(3)$ gaugino soft mass $M_3$, we have extended the original code to make use of the two-loop RGEs generated by SARAH \cite{Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta} and FlexibleSUSY \cite{Athron:2014yba}. The CP-even Higgs masses are calculated including the leading one-loop effective potential contributions given in \cite{Athron:2009bs} and the leading two-loop contributions. To scan over the MSSM parameter space, the equivalent MSSM fine tuning expressions were implemented into a modifed version of SOFTSUSY 3.3.10 \cite{Allanach:2001kg,Allanach:2013kza}. For consistency with the results produced in the $E_6$ models, and for computational speed, for our main scans only the dominant one- and two-loop corrections to the CP-even Higgs masses were included. Finally, in all of the results below the fine tuning was evaluated at the scale $Q=M_{\textrm{SUSY}}=\sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$, where $m_{\tilde{t}_{1,2}}$ are the running $\overline{\textrm{DR}}$ stop masses evaluated at $Q=M_{\textrm{SUSY}}$.

@@

% DH:: I'm not sure if this is worth including? I've put a disclaimer in the
% text, as well.
\begin{figure}[h]
\begin{center}
\resizebox{!}{10cm}{%
\includegraphics{mssm_higgs_tuning.png}}
\caption{Scatter plot of fine tuning in the MSSM as a function of
the lightest Higgs mass, $m_{h_1}$, for the cut-off scales (from 
bottom to top) $M_X=20$ TeV, $M_X=50$ TeV, $M_X=100$ TeV and $M_X=10^{16}$ GeV.}
\label{Fig:mssmhiggstuning}
\end{center}
\end{figure}
To illustrate this, in Fig.~\ref{Fig:mssmstoptuning} we show the
variation in fine tuning when we scan over the stop masses and mixing,
with $500 \textrm{ GeV} \leq m_{Q_3},m_{u_3}\leq 10\textrm{ TeV}$ and
$-3810\textrm{ GeV}\leq A_t\leq -20\textrm{ GeV}$, for $M_X=20$ TeV,
$50$ TeV and $100$ TeV.  The remaining parameters we fix such that at
$M_{SUSY}$ they have the values $\mu=-97.5$ GeV, $B=-84.8$ GeV,
$M_1=92.1$ GeV, $M_2=95.9$ GeV, $M_3=352$ GeV, $A_b=-117.9$ GeV,
$A_\tau=-7.8$ GeV, $m_{L_i}=400$ GeV, $m_{e_i}=204$ GeV,
$m_{Q_{1,2}}=438$ GeV, $m_{u_{1,2}}436$ GeV and $m_{d_i}=438$ GeV
($i=1,2,3$). Although we should stress that making this choice will
lead to a spectrum which is in conflict with the LHC limits, doing so
ensures that fine tuning due to the other parameters is small, so that
we avoid washing out the fine tuning impact of the stops when the
tuning is small\footnote{For models in which the spectrum is heavier,
  when the stop masses are small the fine tuning reaches a lower bound
  imposed by other heavier parameters.} as can be the case when the
stop masses are less than 1 TeV.  Note that the Higgs mass is also
allowed to vary in this scan, as shown in
Fig.~\ref{Fig:mssmhiggstuning}. This reveals the tuning problem which
people have been worrying about since the discovery of the 125 GeV
Higgs boson as we see that raising the stop masses is also pushing up
the Higgs mass, meaning that heavier Higgs masses require more fine
tuning.  However for a low value of the UV scale this tuning is not so
severe unless the stops are very heavy, and a 125 geV Higgs can be
obtained without much tuning in this unrealistic case where we have
minimised other sources of tuning.  On the other hand the tuning
becomes more severe as we increase the cuttoff such that for $M_X =
10^{16}$ GeV a lightest stop masses of $1-3$ TeV can result in a fine
tuning of $\approx 200 - 1000$ as shown in
Fig.~\ref{Fig:mssmstoptuning} and the minimum tuning we see consistent
with a $125$ GeV Higgs is $\approx 200$, as shown in
Fig.~\ref{Fig:mssmhiggstuning}.

@@

Nonetheless the MSSM the impact of chargino limit is shown in
Fig.~\ref{}.  There we see that if the the full parameter space with
$m_{\chi^\pm} < 700$ GeV was excluded the impact would be to make the
tuning in the MSSM with a $20$ TeV cuttoff slightly worse than that of
the E$_6$SSM with the same cuttoff and a $Z^\prime$ mass just larger
than current limits.  would be similar to that can also be attcked
more directly by looking at chargino mass limits as this rasies $\mu$
directly.  In the E$_6$SSM, while raising the chargino limit can have
the same impact in princuiple, due to current limits on the $Z^\prime$
mass already imposing a significant degree of tuning, chargino masses
do not make much of a noticable change.

@@

Finally we want to emphasise that while in Fig.~\ref{Fig:e6ssmvsmssm}
the E$_6$SSM looks more fine tuned than the MSSM this depends on the
cutoff scale.  Indeed in \cite{Athron:2013ipa} a constrained version
of the E$_6$SSM, with the cuttoff at the GUT scale, is considered and
there the cE$_6$SSM was found to be less tuned than the cMSSM. Since a
125 GeV Higgs can be achieved in the E$_6$SSM with ligher stops, then
if the cutoff is large, the larger stop masses of the MSSM can make
that model more fine tuned due to large RGE effects.

As is shown in Fig.~\ref{} the tuning in the MSSM becomes larger
than that of the E$_6$SSM when the cuttoff scale is ?

\section{\label{sec:conclusion}Conclusions}
Prior to stringent experimental constraints on the mass of the
lightest Higgs boson and squarks in supersymmetric models, a simple
picture of a natural SUSY model emerged from theoretcal reasoning,
with soft masses set to similar values at the GUT scale through local
gravity interactions with the hidden sector.  Through the use of
renormalisation group running one can then see that at the EW scale
the stops and soft Higgs masses (which appear directly in the EWSB
conditions) should appear in the the EWSB condition for $M_Z$ and
therefore it was expected that these masses should not be much bigger
than $100$ GeV.  However to disturb this elegant picture first LEP
placed constraints on the Higgs mass, requiring it to be above 114.4
GeV \cite{}, which already introduced significant tuning for
constrained models, and then the LHC meaured the Higgs to be around
$125$ GeV.

However

@@

Provided that one does not run over too large a range of scales, the solutions to the
RG equations for a model can be reasonably well approximated by a Taylor series, Eq. (\ref{eq:rgeapproxsoln}). 
For a parameter $p$, this reads

@@

We have constructed the necessary series solutions in both the MSSM and the $U(1)$ extended models. Due to the smallness of the first and second generation Yukawa couplings, we neglect them in our calculations. The corresponding soft SUSY breaking 
trilinears are likewise omitted. Additionally, all soft mass matrices are assumed diagonal, and the gaugino masses are taken to be real.\\ \\
In the MSSM, the relevant parameters for the fine tuning calculation are $\mu$, $B$, $m_{H_u}^2$, $m_{H_d}^2$ at tree level. The one- and two-loop RGEs for these parameters may be found in \cite{Martin:1993zk}. The corresponding $O(t^2)$ contributions are

@@

If in addition the one-loop contributions to the effective potential from top and stop loops are included, it is also
necessary to construct the expansions for $m_{Q3}^2$, $m_{u3}^2$ and $A_t$. The coefficients read

@@

Similarly we can obtain the two-loop $\beta$ functions and coefficients $b_p^{(2)}$ for a general set of $U(1)'$ charges. Two-loop RGEs for the gauge and Yukawa couplings, gaugino masses and soft trilinears, along with the one-loop RGEs for the soft scalar masses, were given in \cite{Athron:2009bs} for the particular case of the E$_6$SSM. At tree level in the EWSB conditions the parameters that must be considered are $\lambda$, $A_\lambda$, $m_{H_u}^2$, $m_{H_d}^2$, $m_S^2$ and $g_1$, $g_2$ and $g_1'$. The one- and two-loop contributions to the $\beta$ function for $\lambda$ are

>>>>

V_{E_6}^F&=\lambda^2|S|^2(|H_1|^2+|H_2|^2)+\lambda^2|(H_1\cdot
H_2)|^2,\label{eq:E6VFterms}\\ V_{E_6}^D&=\frac{\bar{g}^2}{8}\left (
|H_2|^2-|H_1|^2\right )^2+\frac{g_2^2}{2}|H_1^\dagger
H_2|^2+\frac{g_1'^2}{2}(\tilde{Q}_1|H_1|^2+\tilde{Q}_2|H_2|^2+\tilde{Q}_S|S|^2)^2,\label{eq:E6VDterms}\\ V_{E_6}^{\textrm{soft}}&=m_S^2|S|^2+m_{H_d}^2|H_1|^2+m_{H_u}^2|H_2|^2+\Big
[\lambda A_\lambda S(H_1\cdot H_2)+\textrm{h.c.}\Big
].\label{eq:E6Vsoft}

@@

The quantities $D_{H_1}$, $D_{H_2}$ and $D_S$ appearing above are
$U(1)'$ $D$-term contributions that are absent in the MSSM and NMSSM
and are given by

@@

D_\phi\equiv \frac{g_1'^2}{2}\left (
\tilde{Q}_1v_1^2+\tilde{Q}_2v_2^2+\tilde{Q}_Ss^2\right
)\tilde{Q}_\phi.

@@

As was noted in \cite{Athron:2013ipa}, the first two of these
conditions may be rewritten in the form

@@

with $M_Z^2=\bar{g}^2v^2/4$, $v^2=v_1^2+v_2^2$ and $\tan\beta =
v_2/v_1$ and where we have for convenience absorbed the effects of the
loop corrections into the soft masses,

@@

\tilde{m}_{H_d}^2&=m_{H_d}^2+\frac{1}{v_1}\frac{\partial \Delta
  V_{E_6}}{\partial
  v_1},\\ \tilde{m}_{H_u}^2&=m_{H_u}^2+\frac{1}{v_2}\frac{\partial
  \Delta V_{E_6}}{\partial v_2}.

@@

Written like so it is evident that the fine tuning contribution coming
from the new $D$-terms depends both on the $U(1)'$ charges and the
$Z^\prime$ mass, and that the tuning can be expected to increase with
$M_{Z^\prime}$. As shall be shown below, the exact size of this tuning
then depends strongly on the choice of $U(1)'$ charges, via the
coefficient $d$.

@@

\Delta_p=\left | \frac{\partial \log M_Z^2}{\partial \log p}\right
|=\left | \frac{p}{M_Z^2}\frac{\partial M_Z^2}{\partial p}\right |,

@@

where $M_Z^2=\bar{g}^2v^2/4$. The coefficient $\Delta_p$ measures the
fractional variation in $M_Z^2$ resulting from a given variation in
the parameter $p$. The overall fine tuning is then taken to be $\Delta
= \underset{i}{\max}\,\{\Delta_{p_i}\}$.\\ \\

The sensitivity coefficients $\Delta_p$ may be calculated directly
from the expression for $M_Z^2$ in terms of the $p_i$ for a particular
model, which leads to a so-called master formula for calculating the
fine tuning. A master formula for the E$_6$SSM, obtained from the tree
level scalar potential, was presented in \cite{Athron:2013ipa}. In
order to derive the expression presented there, the fact that $s \gg
v$ was made use of to neglect certain $O(v^2)$ terms in the EWSB
conditions, greatly simplifying the final result. For the purposes of
exploring a wider class of $E_6$ inspired models, we have rederived
the master formula without neglecting any $O(v^2)$ terms. The more
complete tree-level master formula is somewhat complicated. This is
because, unlike in the MSSM, even at tree level it is not possible to
solve explicitly for the VEVs $v_1$, $v_2$ in terms of the Lagrangian
parameters. It may be written in the form \be \Delta_p=|C|^{-1}\times
\frac{|p|}{M_Z}\bigg|\sum_{q}\tilde{\Delta}_q\frac{\partial
  q}{\partial p}\bigg|,

@@

\ee where the sum is over all low energy running parameters appearing
in the tree level EWSB conditions, i.e. $q\in\{\lambda, A_\lambda,
m_{H_d}^2,m_{H_u}^2,m_S^2,g_1,g_2,g_1'\}$. Expressions for the
quantity $C$ and the $\tilde{\Delta}_q$ appearing above are given in
Appendix \ref{app:masterformula}. It should be noted that the effects
of $U(1)$ mixing are neglected in deriving
Eq. (\ref{eq:E6masterformula}).\\ \\ However, it is well known in the
MSSM that radiative corrections can significantly change (indeed,
reduce) the fine tuning \cite{Cassel:2010px}. It is therefore
important when studying the fine tuning to include loop corrections to
the effective potential in the fine tuning measure. To do so it is
most convenient to work in terms of the EWSB conditions
Eq. (\ref{eq:E6EWSBConditions}), rather than
Eq. (\ref{eq:E6MZequation}). The general procedure is as follows
\cite{Ellwanger:2011mu}. For a model in which $m$ fields develop real
VEVs (e.g. $m=2$ in the MSSM, $m=3$ in the NMSSM and in the $E_6$
models considered), we require that the $m$ minimization conditions

@@

continue to hold under an arbitrary variation in a model parameter
$p\rightarrow p+\delta p$, so that the variations $\delta f_i$ satisfy

@@

Each $f_i$ is a function of the VEVs $v_j$ and $l$ running parameters
$q_k$ evaluated at the scale of EWSB, $f_i=f_i(v_j,q_k)$. Thus for
each $f_i$ we have that

@@

\delta f_i = \sum_{j=1}^m \frac{\partial f_i}{\partial
  v_j}\frac{\partial v_j}{\partial p}+\sum_{k=1}^l \frac{\partial
  f_i}{\partial q_k}\frac{\partial q_k}{\partial p}=0.

@@

The quantities $\frac{\partial f_i}{\partial v_j}$ are the elements of
the CP-even Higgs squared mass matrix $M_h^2$ of the model before
rotating into the mass eigenstate basis. When evaluated for all $n$
model parameters, the above system of equations can be concisely
expressed as

@@

The quantities forming the first matrix on the right-hand side, along
with $M_h^2$, are easily calculated analytically. The remaining
derivatives $\partial q_k /\partial p$ must be determined using the
RGEs. Once these have been obtained, it is straightforward to solve
for the $\partial v_i /\partial p$. The sensitivity coefficients
$\Delta_p$ are then simply linear combinations of the $\partial
v_i/\partial p$ and $\partial q_k/\partial p$. Radiative corrections
may be easily incorporated by including the Coleman-Weinberg potential
contributions in the EWSB conditions.\\ \\ Evaluating the derivatives
$\partial q_k/\partial p$ must (in general) be done by numerically
integrating the two-loop RGEs. This is time consuming and presents an
obstacle to doing large scans of the parameter space. For studying
models defined at low energies, as we do here, we can take advantage
of the fact that the amount of running involved is small. This makes
it possible to use approximate analytic solutions to the RGEs that
exhibit good accuracy over the range of scales considered. Given the
two-loop RG equation for a parameter $q$,

@@

a Taylor series expansion of the solution may be used to obtain the
parameter at the scale $Q$,

@@

Expanded to this order we obtain the leading log (LL) and next-to-LL
(NLL) contributions at two-loop order. The $\mathcal{O}(t^2)$ terms
not displayed above are formally of 3-loop order and are
neglected. The derivative of the one-loop $\beta$ function is given by

@@

where the sum is over all running parameters appearing in
$\beta_q^{(1)}$. The $\beta$ functions appearing on the right hand
side of Eqs. (\ref{eq:rgeapproxsoln}) and
(\ref{eq:1loopbetaderivative}) are evaluated at the input scale $M_X$,
giving a simple analytic expression for the parameters at the scale of
EWSB in terms of the model parameters at $M_X$. Explicit results for
the relevant series expansions in the MSSM and $E_6$ models are
presented in Appendix \ref{app:rges}.

@@

Using the approach outlined above, we are able to scan the low energy
parameter space of the MSSM and E$_6$SSM and calculate the fine tuning
in each. To do so, we implemented the above expressions for computing
the fine tuning in a modified version of the E$_6$SSM spectrum
generator that was used in \cite{Athron:2013ipa}. This code
implemented two-loop RGEs for all parameters except the soft scalar
masses. In order to properly include the fine tuning impact of the
$SU(3)$ gaugino soft mass $M_3$, we have extended the original code to
make use of the two-loop RGEs generated by SARAH
\cite{Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta} and
FlexibleSUSY \cite{Athron:2014yba}. The CP-even Higgs masses are
calculated including the leading one-loop effective potential
contributions given in \cite{Athron:2009bs} and the leading two-loop
contributions. To scan over the MSSM parameter space, the equivalent
MSSM fine tuning expressions were implemented into a modifed version
of SOFTSUSY 3.3.10 \cite{Allanach:2001kg,Allanach:2013kza}. For
consistency with the results produced in the $E_6$ models, and for
computational speed, for our main scans only the dominant one- and
two-loop corrections to the CP-even Higgs masses were
included. Finally, in all of the results below the fine tuning was
evaluated at the scale
$Q=M_{\textrm{SUSY}}=\sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$, where
$m_{\tilde{t}_{1,2}}$ are the running $\overline{\textrm{DR}}$ stop
masses evaluated at $Q=M_{\textrm{SUSY}}$.

@@

% DH:: I'm not sure if this is worth including? I've put a disclaimer in the
% text, as well.
\begin{figure}[h]
\begin{center}
\resizebox{!}{10cm}{%
\includegraphics{mssm_higgs_tuning.png}}
\caption{Scatter plot of fine tuning in the MSSM as a function of
the lightest Higgs mass, $m_{h_1}$, for the cut-off scales (from 
bottom to top) $M_X=20$ TeV, $M_X=50$ TeV, $M_X=100$ TeV and $M_X=10^{16}$ GeV.}
\label{Fig:mssmhiggstuning}
\end{center}
\end{figure}
To illustrate this, in Fig.~\ref{Fig:mssmstoptuning} we show the
variation in fine tuning when we scan over the stop masses and mixing,
with $500 \textrm{ GeV} \leq m_{Q_3},m_{u_3}\leq 10\textrm{ TeV}$ and
$-3810\textrm{ GeV}\leq A_t\leq -20\textrm{ GeV}$, for $M_X=20$ TeV,
$50$ TeV and $100$ TeV.  The remaining parameters we fix such that at
$M_{SUSY}$ they have the values $\mu=-97.5$ GeV, $B=-84.8$ GeV,
$M_1=92.1$ GeV, $M_2=95.9$ GeV, $M_3=352$ GeV, $A_b=-117.9$ GeV,
$A_\tau=-7.8$ GeV, $m_{L_i}=400$ GeV, $m_{e_i}=204$ GeV,
$m_{Q_{1,2}}=438$ GeV, $m_{u_{1,2}}436$ GeV and $m_{d_i}=438$ GeV
($i=1,2,3$). Although we should stress that making this choice will
lead to a spectrum which is in conflict with the LHC limits, doing so
ensures that fine tuning due to the other parameters is small, so that
we avoid washing out the fine tuning impact of the stops when the
tuning is small\footnote{For models in which the spectrum is heavier,
  when the stop masses are small the fine tuning reaches a lower bound
  imposed by other heavier parameters.} as can be the case when the
stop masses are less than 1 TeV.  Note that the Higgs mass is also
allowed to vary in this scan, as shown in
Fig.~\ref{Fig:mssmhiggstuning}. This reveals the tuning problem which
people have been worrying about since the discovery of the 125 GeV
Higgs boson as we see that raising the stop masses is also pushing up
the Higgs mass, meaning that heavier Higgs masses require more fine
tuning.  However for a low value of the UV scale this tuning is not so
severe unless the stops are very heavy, and a 125 geV Higgs can be
obtained without much tuning in this unrealistic case where we have
minimised other sources of tuning.  On the other hand the tuning
becomes more severe as we increase the cutoff such that for $M_X =
10^{16}$ GeV a lightest stop masses of $1-3$ TeV can result in a fine
tuning of $\approx 200 - 1000$ as shown in
Fig.~\ref{Fig:mssmstoptuning} and the minimum tuning we see consistent
with a $125$ GeV Higgs is $\approx 200$, as shown in
Fig.~\ref{Fig:mssmhiggstuning}.

@@

Nonetheless the MSSM the impact of chargino limit is shown in
Fig.~\ref{}.  There we see that if the the full parameter space with
$m_{\chi^\pm} < 700$ GeV was excluded the impact would be to make the
tuning in the MSSM with a $20$ TeV cutoff slightly worse than that of
the E$_6$SSM with the same cutoff and a $Z^\prime$ mass just larger
than current limits.  would be similar to that can also be attcked
more directly by looking at chargino mass limits as this rasies $\mu$
directly.  In the E$_6$SSM, while raising the chargino limit can have
the same impact in princuiple, due to current limits on the $Z^\prime$
mass already imposing a significant degree of tuning, chargino masses
do not make much of a noticable change.

@@

Finally we want to emphasise that while in Fig.~\ref{Fig:e6ssmvsmssm}
the E$_6$SSM looks more fine tuned than the MSSM this depends on the
cutoff scale.  Indeed in \cite{Athron:2013ipa} a constrained version
of the E$_6$SSM, with the cutoff at the GUT scale, is considered and
there the cE$_6$SSM was found to be less tuned than the cMSSM. Since a
125 GeV Higgs can be achieved in the E$_6$SSM with ligher stops, then
if the cutoff is large, the larger stop masses of the MSSM can make
that model more fine tuned due to large RGE effects.

As is shown in Fig.~\ref{} the tuning in the MSSM becomes larger
than that of the E$_6$SSM when the cutoff scale is ?

\section{\label{sec:conclusion}Conclusions}

Prior to stringent experimental constraints on the mass of the
lightest Higgs boson and squarks in supersymmetric models, a simple
picture of a natural SUSY model emerged from theoretical reasoning,
with soft masses set to similar values at the GUT scale through local
gravity interactions with the hidden sector.  Through the use of
renormalisation group running one can then see that at the EW scale
the stops and soft Higgs masses (which appear directly in the EWSB
conditions) should appear in the the EWSB condition for $M_Z$ and
therefore it was expected that these masses should not be much bigger
than $100$ GeV.  However to disturb this elegant picture first LEP
placed constraints on the Higgs mass, requiring it to be above 114.4
GeV \cite{}, which already introduced significant tuning for
constrained models,and then the LHC measured the Higgs to be around
$125$ GeV, since heavy stops are required to raised the lightest
Higgs mass above it's tree level bound.

$U(1)$ extensions motivated by the $\mu$-problem, $E_6$ GUT theories
and the connection to string theory contain both $F-$ and $D-$ term
contributions to the light Higgs mass which can raise the tree level
mass, evading the need for large radiative corrections to increase it.
However such models come with their own fine tuning problem, where the
$Z^\prime$ mass appears in the EWSB condition for $M_Z$ at tree level.
While in a previous study of the constrained E$_6$SSM it was found
that the tuning is less severe than the MSSM, it was still significant.

In light of such difficulties it is worth considering whether the
simple picture which emerged is wrong in some way and if there are
other possibilities that allow naturalness.  Or to phrase this in a
more challenging manner are there ways to constrain the naturalness of
these models that do not rely upon assumptions about how SUSY is
broken?

We have investigated this question here in the context of the MSSM and
U(1) extensions.  Since the RG evolution links the soft masses
together and causes these problems from stop and gluino masses the
most conservative approach to placing naturalness limits is to choose
a low cutoff. We find that in the MSSM the most direct way to
constrain naturalness of the model, without making assumptions about
the SUSY breaking scale is through limits on the chargino
masses. Current LHC limits on charginos are not model independent and
thereby leave many gaps where on can have light charginos.

In contrast we find that in $U(1)$ extensions of the MSSM there is an
additional way to constrain the naturalness of the models, which is
through the $Z^\prime$ mass limit.  We find when we impose a low
cutoff of $20$ TeV for setting the soft masses, the lowest tuning in
the E$_6$SSM is compatible with a $Z^\prime$ mass of $2.5$ TeV was
$\Delta \approx 100$, while if the LHC run II can place a limit of
$4.5$ TeV on $M_Z^\prime$ then the tuning would be just under $400$.
By comparison the current situation in the MSSM only requires a tuning
of around $20-30$. [Dylan: update numbers a bit -- I suck at reading
  them off the graphs:)] This should be interpreted as saying that in
the most conservative limits one can place on naturalness on these
models, the tuning in the E$_6$SSM is worse.  However if the chargino
limits blah blah blah then the situation in the two models would be
the same.

This should also be contrasted with what happens as we raise the
cutoff scale.  We showed that when the cutoff reaches $M_X =
[insert]$ the tuning of the MSSM becomes larger than that of the
E$_6$SSM.  Therefore the conclusion is not that one model is less
tuned than another, but rather that the answer depends on what
assumptions are made about the SUSY breaking.

Finally we also looked at the tuning in different $U(1)$ extensions for
fixed $\tan \beta = 10$.  We found that in all cases the fine tuning
was much worse for the larger $Z^\prime$ mass, further emphasising
the importance they play in these $U(1)$ extensions.  Interestingly
the actual level in tuning varied quite a bit between these models
with the $U(1)_I$ model showing the least tuning due to the vanishing
charge of the $H_u$ state. This model is quite interesting in the
sense that it provides a solution to the $\mu$-problem while avoiding
the large tuning (with current limits) from the $Z^\prime$
mass. However one should remember we are looking at conservative limits
on naturalness here and there is no solution to the usual tuning from
the large stops needed to get a $125$ GeV Higgs in this model.  In any
case future LHC limits on the $Z^\prime$ mass would be especially
interesting for this case as they would significantly raise the
minimum level of tuning which can be obtained.

@@

Provided that one does not run over too large a range of scales, the
solutions to the RG equations for a model can be reasonably well
approximated by a Taylor series, Eq. (\ref{eq:rgeapproxsoln}).  For a
parameter $p$, this reads

@@

We have constructed the necessary series solutions in both the MSSM
and the $U(1)$ extended models. Due to the smallness of the first and
second generation Yukawa couplings, we neglect them in our
calculations. The corresponding soft SUSY breaking trilinears are
likewise omitted. Additionally, all soft mass matrices are assumed
diagonal, and the gaugino masses are taken to be real.\\ \\ In the
MSSM, the relevant parameters for the fine tuning calculation are
$\mu$, $B$, $m_{H_u}^2$, $m_{H_d}^2$ at tree level. The one- and
two-loop RGEs for these parameters may be found in
\cite{Martin:1993zk}. The corresponding $O(t^2)$ contributions are

@@

If in addition the one-loop contributions to the effective potential
from top and stop loops are included, it is also necessary to
construct the expansions for $m_{Q3}^2$, $m_{u3}^2$ and $A_t$. The
coefficients read

@@

Similarly we can obtain the two-loop $\beta$ functions and
coefficients $b_p^{(2)}$ for a general set of $U(1)'$
charges. Two-loop RGEs for the gauge and Yukawa couplings, gaugino
masses and soft trilinears, along with the one-loop RGEs for the soft
scalar masses, were given in \cite{Athron:2009bs} for the particular
case of the E$_6$SSM. At tree level in the EWSB conditions the
parameters that must be considered are $\lambda$, $A_\lambda$,
$m_{H_u}^2$, $m_{H_d}^2$, $m_S^2$ and $g_1$, $g_2$ and $g_1'$. The
one- and two-loop contributions to the $\beta$ function for $\lambda$
are
